A python toolbox to tackle the curse of imbalanced datasets in machine learning

### 2016-06-18 main.tex
Advance project vision

\begin{abstract}
\emph{UnbalancedDataset} is an open-source python toolbox aiming at providing a wide range of methods to cope with the problem of imbalanced dataset frequently encountered in machine learning and pattern recognition.
The state-of-the-art methods implemented can be categorized into 4 different sampling strategies: (i) under-sampling, (ii) over-sampling, (iii) combination of over- and under-sampling, and (iv) ensemble learning methods.
The proposed toolbox only depends of \emph{numpy}, \emph{scipy}, and \emph{scikit-learn} and is distributed under MIT license.
The implementation pattern is identical to the \emph{scikit-learn} API to offer compatibility.
Documentation, unit tests as well as integration tests are provided to allow new developers to contribute.
The toolbox is publicly available in GitHub \url{https://github.com/fmfn/UnbalancedDataset} as well as the documentation \url{https://github.io}.
\end{abstract}

@@

\begin{description}
  \item[Quality insurance] In order to ensure code quality, a set of unit tests is provided leading to a coverage of 99 \% for the release 0.1 of the toolbox. Furthermore, the code quality is ensured by following \texttt{PEP8} standard and each contribution is automatically check through landscape.
  \item[Continuous integration]
  \item[Documentation]
\end{description}

In order to be sustainable and improved over years, the current toolbox offer a detailed documentation as well as a unit and integration tests.
A set of \emph{getting started} guide, examples, and API documentation is available.
The API documentation is automatically generated using \emph{sphinx} and \emph{numpydoc}.
A set of unit tests are implemented to allow back compatibility and integration of new pull-requests.
Furthermore, automatic continuous integration is also available through Travis CI.
Bugs can be reported through the issue tracker system provided by GitHub.

@@

The implementation rely on \emph{numpy}, \emph{scipy}, and \emph{scikit-learn}.
The implementation pattern is identical to the \emph{scikit-learn} API to offer compatibility.
Each class implements 3 main functions inspired from the \emph{scikit-learn} API:
(i) \emph{fit} computes the parameter values which are later needed to transform the data into a balanced set;
(ii) \emph{transform} performs the sampling and return the data with the desired balancing ratio;
and (iii) \emph{fit\_transform} is equivalent of calling the function \emph{fit} follow the function \emph{transform}.

@@

Three strategies can be employed to overcome the problem of imbalanced datasets: 
(i) under-sampling, (ii) over-sampling, (iii) a combination of both, and (iv) ensemble learning.
The following sections give an overview of the techniques used to tackle this issue.

@@

Considering the problem of imbalanced, under-sampling is performed such that the number of samples of the majority class is reduced to be equal to the number of samples of the minority class.
The following methods are considered to perform such balancing.

\begin{description}
  \item[Random under-sampling] is performed by randomly selecting without replacement a subset of samples from the majority class such that the number of samples is then equal in both minority and majority classes.
  \item[Tomek links] can be used to under-sample the majority class of the original dataset~\cite{tomek1976two}.
Let define a pair of nearest neighbour samples $(x_i, x_j)$ such that their associated class label $y_i \neq y_j$.
The pair $(x_i, x_j)$ is defined as a Tomek link if, by relaxing the class label differentiation constraint, there is no other sample $x_k$ defined as the nearest neighbour of either $x_i$ or $x_j$.
Under-sampling is performed by removing the samples belonging to the majority class and forming a Tomek link.
It can be noted that this under-sampling strategy does not enforce a strict balancing between the majority and the minority classes.
  \item[Cluster centroids method] refers to the use of a $k$-means to cluster the feature space such that $k$ is set to be equal to the number of samples composing the minority class.
Hence, the centroids of these clusters define the new samples of the majority class. 
  \item[NearMiss] offers three different methods to under-sample the majority class~\cite{mani2003knn}.
In NearMiss-1, samples from the majority class are selected such that for each sample, the average distance to the $k$ nearest neighbour samples from the minority class is minimum.
NearMiss-2 diverges from NearMiss-1 by considering the $k$ farthest neighbours samples from the minority class.
In NearMiss-3, a subset $M$ containing samples from the majority class is generated by finding the $m$ nearest neighbours from each sample of the minority class.
Then, samples from the subset $M$ are selected such that for each sample, the average distance to the $k$ nearest neighbour samples from the minority class is maximum.
In our experiment, $k$ and $m$ are fixed to 3.
  \item[Neighbourhood cleaning rule] consists of applying two rules depending on the class of each sample~\cite{laurikkala2001improving}.
Let define $x_i$ as a sample of the dataset with its associated class label $y_i$.
Let define $y_m$ as the class of the majority vote of the $k$ nearest neighbours of the sample $x_i$.
If $y_i$ corresponds to the majority class and $y_i \neq y_m$, $x_i$ is rejected from the final subset.
If $y_i$ corresponds to the minority class and and $y_i \neq y_m$, then the $k$ nearest neighbours are rejected from the final subset.
In our experiment $k$ is fixed to 3.
\end{description}

@@

\acks{We would like to acknowledge support for this project
from GitHub, Travis CI, and Gitter.}

>>>>

\begin{abstract}
\emph{UnbalancedDataset} is an open-source python toolbox aiming at providing a wide range of methods to cope with the problem of imbalanced dataset frequently encountered in machine learning and pattern recognition.
The state-of-the-art methods implemented can be categorized into 4 different sampling strategies: (i) under-sampling, (ii) over-sampling, (iii) combination of over- and under-sampling, and (iv) ensemble learning methods.
The proposed toolbox only depends of \emph{numpy}, \emph{scipy}, and \emph{scikit-learn} and is distributed under MIT license.
The implementation pattern is identical to the \emph{scikit-learn} API to offer compatibility.
Documentation, unit tests as well as integration tests are provided to allow new developers to contribute.
The toolbox is publicly available in GitHub \url{https://github.com/fmfn/UnbalancedDataset} as well as the documentation \url{http://fmfn.github.io/UnbalancedDataset/}.
\end{abstract}

@@

\begin{description}
  \item[Quality insurance] In order to ensure code quality, a set of unit tests is provided leading to a coverage of 99 \% for the release 0.1 of the toolbox.
Furthermore, the code consistency is ensured by following \texttt{PEP8} standards and each new contribution is automatically check through landscape, which provides metrics related to code quality.
  \item[Continuous integration] To allow user and developer to either use or contribute this toolbox, Travis CI is used to easily integrate new code and ensure back-compatibility.
  \item[Community-based development] All the development is performed in a collaborative manner.
Tools as git, GitHub, and gitter are used to ease collaborative programming, issue tracking, code integration, and ideas discussions.
  \item[Documentation] A consistent API documentation is provided using \emph{sphinx} and \emph{numpydoc}.
Additional installation guide, examples, and tutorial are also provided and centralized on GitHub\footnote{\url{http://fmfn.github.io/UnbalancedDataset/}}.
\end{description}

@@

The implementation rely on \emph{numpy}, \emph{scipy}, and \emph{scikit-learn}.
Each class implements 3 main functions inspired from the \emph{scikit-learn} API:
(i) \emph{fit} computes the parameter values which are later needed to transform the data into a balanced set;
(ii) \emph{transform} performs the sampling and return the data with the desired balancing ratio;
and (iii) \emph{fit\_transform} is equivalent of calling the function \emph{fit} follow the function \emph{transform}.

@@

The \textbf{UnbalancedDataset} toolbox provides four different strategies to tackle the problem of imbalanced dataset:
(i) under-sampling, (ii) over-sampling, (iii) a combination of both, and (iv) ensemble learning.
The following sections give an overview of the techniques implemented.

@@

Considering the problem of imbalanced, under-sampling is performed such that the number of samples of the majority class is reduced.
The following methods are considered to perform such balancing.

\begin{description}
  \item[Random under-sampling] is performed by randomly selecting without replacement a subset of samples from the majority class such that the number of samples is then equal in both minority and majority classes.
  \item[Cluster centroids method] refers to the use of a $k$-means to cluster the feature space such that $k$ is set to be equal to the number of samples composing the minority class.
Hence, the centroids of these clusters define the new samples of the majority class. 
  \item[Condensed nearest neighbours] Something here.
  \item[Edited nearest neighbours] Something here.
  \item[Instance hardness neigbours]
  \item[NearMiss] offers three different methods to under-sample the majority class~\cite{mani2003knn}.
In NearMiss-1, samples from the majority class are selected such that for each sample, the average distance to the $k$ nearest neighbour samples from the minority class is minimum.
NearMiss-2 diverges from NearMiss-1 by considering the $k$ farthest neighbours samples from the minority class.
In NearMiss-3, a subset $M$ containing samples from the majority class is generated by finding the $m$ nearest neighbours from each sample of the minority class.
Then, samples from the subset $M$ are selected such that for each sample, the average distance to the $k$ nearest neighbour samples from the minority class is maximum.
In our experiment, $k$ and $m$ are fixed to 3.
  \item[One-sided selection] Something here.
  \item[Neighbourhood cleaning rule] consists of applying two rules depending on the class of each sample~\cite{laurikkala2001improving}.
Let define $x_i$ as a sample of the dataset with its associated class label $y_i$.
Let define $y_m$ as the class of the majority vote of the $k$ nearest neighbours of the sample $x_i$.
If $y_i$ corresponds to the majority class and $y_i \neq y_m$, $x_i$ is rejected from the final subset.
If $y_i$ corresponds to the minority class and and $y_i \neq y_m$, then the $k$ nearest neighbours are rejected from the final subset.
In our experiment $k$ is fixed to 3.
  \item[Tomek links] can be used to under-sample the majority class of the original dataset~\cite{tomek1976two}.
Let define a pair of nearest neighbour samples $(x_i, x_j)$ such that their associated class label $y_i \neq y_j$.
The pair $(x_i, x_j)$ is defined as a Tomek link if, by relaxing the class label differentiation constraint, there is no other sample $x_k$ defined as the nearest neighbour of either $x_i$ or $x_j$.
Under-sampling is performed by removing the samples belonging to the majority class and forming a Tomek link.
It can be noted that this under-sampling strategy does not enforce a strict balancing between the majority and the minority classes.

\end{description}

@@

\acks{We would like to acknowledge support for this project
from git, GitHub, Travis CI, and Gitter.}
